---
title: "Crime Rates"
---

Follow along with the demo to learn how to calculate crime rates using R and the Tidyverse. We'll also introduce a new library, janitor, which helps make column names easier to work with. You should get started by loading the tidyverse in a new R block, then install and load janitor.

### Task 1: Load the tidyverse and install and load janitor
```{r}
library(tidyverse)
install.packages("janitor")
library("janitor")
```


### Task 2: Load the crime dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv
```{r}
crimes <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv") |> clean_names()
```



### Task 3: Use janitor to clean up the column names



### Task 4: Calculate the total crime rate for each year (picking our unit!)

```{r}
crimes |> mutate(total_rate= grand_total/population * 10000)
```


### Task 5: Calculate the violent and property crime rates for each year (same unit)
```{r}
crimes |> mutate(violent_rate= violent_crime_total/population * 10000,property_rate= property_crime_total/population * 10000)
```



### Task 6: Describe our results

Write a few sentences about what you found in the data. What years had the highest crime rates? What years had the lowest crime rates? Does the pattern match your expectations?

Marylanders saw the lowest ever violent crimes rates going into the new decade. 

The violent crime rate reduced to the lowest its ever been in 2020 with 41 violent crimes per 10,000 people. Maryland's highest crime rates were in 1992 at around 99 crimes per 10,000 people.

Property crimes are a similar yet different story. They were also at an all-time low in 2020. However... Maryland saw 162 property crimes per 10,000 people that year. The highest year for property crimes in Maryland was 1995. There were 530 property crimes per 10,000 people. 




When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.

