---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse


```{r}
library(tidyverse)
```

```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```




### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv



### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`

```{r}
umd_fees |> mutate(pct_change=(`Fall 2024`- `Fall 2021`)/`Fall 2021` * 100)
```
```{r}
umd_fees_pct <- umd_fees |>  mutate(pct_change=(`Fall 2024`- `Fall 2021`)/`Fall 2021` * 100)

```



### Task 4: Describe our results

Write a few sentences about what you found in the data. What's the story here?

University of Maryland students' bill is steeper over the past three years. A relatively-new category called Student Sustainability and the university's bus system are costing students more. !



When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
