---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

The number of reported "accidents" steadily increased throughout the year.However homicides very important. The categories are described weird like what is the difference between accident and accident with impound? Also what is the difference between non-residential robbery and theft? 

### Task 4: Calculating Averages

Calculate the total number of incidents for each category using `mutate()`, calling the new column `total_incidents Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
library(tidyverse)
'ACCIDENT' <- 435+430+429+488+529+463+462+497+465+569+500+509
'ACCIDENT WITH IMPOUND' <-316+283+299+303+353+334+315+313+347+331+310+345
'ASSAULT' <- 71+58+71+74+86+75+62+70+75+73+73+63
'ASSAULT  SHOOTING' <- sum(13,10,12,20,15,21,21,14,21,16,20)
'ASSAULT  WEAPON' <- sum(23,18,12,20,8,14,16,14,16,10,10,13)
'AUTO  STOLEN'<- sum(393, 412,335,279,394,442,421,420,426,462,420,386)
'AUTO  STOLEN & RECOVERED' <- sum(36,39,23,38,48,48,58,37,36,44,33,47)
'B & E  COMMERCIAL' <- 36+31+29+30+48+48+41+56+37+53+52+75
'B & E  OTHER' <- 5+7+6+11+10+4+5+6+6+10+9+6
'B & E  RESIDENTIAL' <- 38+35+27+52+41+42+30+27+32+46+39+35
'HOMICIDE' <- 5+4+7+5+6+8+7+7+5+3+7+6
'ROBBERY  COMMERCIAL' <- 25+33+19+26+23+17+20+29+14+12+12+12
'ROBBERY  OTHER' <- 62+52+49+52+42+63+36+46+70+57+71+59
'ROBBERY  RESIDENTIAL' <- 5+1+4+2+3+3+3+2+2+3
'ROBBERY  VEHICLE' <- 23+42+16+21+31+31+41+21+15+21+24+13
'SEX OFFENSE' <- 13+5+14+9+11+14+15+10+13+16+14+13
'THEFT' <- 305+253+258+262+285+280+279+260+281+363+293+290
'THEFT FROM AUTO' <- 429+322+348+323+375+395+449+408+388+407+450+409

'total_incidents' <- 5776+3849+851+183+174+4790+487+536+85+444+70+242+659+28+299+147+3409+4703
'average_incidents' <- 26732/12
```

```



### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

Theft, accident, accident with impound, theft from auto have the highest rates. 


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
'august_change' <-2220
'july_change' <- 2182
'june_change' <- 2252

pg_crime |> mutate(pct_change=(`june_change`- `august_change`)/`august change` * 100)



### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?
R won't run
### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Crime has steadily increased and Prince George's County across last year. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
